---
title: "ravenplot_examples"
author: "Diana Tereshchenko"
date: "2026-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(patchwork)
library(ravenplot)
load_raven_fonts()
```

# Introduction 

The `ravenplot` package extends ggplot2 by providing:

- 10 color palettes: "vampire", "vampire_heatmap", "chocolate", "chocolate_heatmap", "witchy", "witchy_heatmap", "lotus", "lotus_heatmap", "tealish", "tealish_heatmap" 

``` {r, echo=FALSE}
plot_palette <- function(palette_name) {

  colors <- get_raven_palette(palette_name)

  df <- data.frame(
    x = seq_along(colors),
    y = 1,
    col = colors
  )

  ggplot(df, aes(x, y, fill = col)) +
    geom_tile() +
    scale_fill_identity() +
    labs(
      title = paste("Palette:", palette_name)
    ) +
    theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5)
    )
}

p1_1 <- plot_palette("vampire")
p1_2 <- plot_palette("vampire_heatmap")

p2_1 <- plot_palette("witchy")
p2_2 <- plot_palette("witchy_heatmap")

p3_1 <- plot_palette("tealish")
p3_2 <- plot_palette("tealish_heatmap")

p4_1 <- plot_palette("lotus")
p4_2 <- plot_palette ("lotus_heatmap")

p5_1 <- plot_palette("chocolate")
p5_2 <- plot_palette("chocolate_heatmap")

(p1_1 | p1_2) / (p2_1 | p2_2) / (p3_1 | p3_2) / (p4_1 | p4_2) / (p5_1 | p5_2)



```

- 7 point styles (Unicode-based symbols)

- 4 font themes: "fancy", "elegant", "simple", "standard"
Bar plot

- 2 background themes

# Scatter plot

``` {r}
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_raven_point(dot_type = "heart", size = 5) +
  scale_color_manual(values = get_raven_palette("vampire")) +
  labs(
    title = "Scatter Plot",
    subtitle = "Raven points with vampire palette",
    x = "Weight",
    y = "MPG"
  ) +
  theme_raven(
    background_type = "vintage",
    font_theme = "fancy",
    color_theme = "vampire"
  ) 

```

# Bar plot 

```{r}

ggplot(mpg, aes(class, fill = class)) +
  geom_bar() +
  scale_fill_manual(values = get_raven_palette("vampire")) +
  labs(
    title = "Bar Plot",
    subtitle = "Vehicle classes",
    x = "Class",
    y = "Count"
  ) +
  theme_raven(
    background_type = "vintage",
    font_theme = "elegant",
    color_theme = "vampire",
    major_grid = FALSE
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Line plot 

``` {r}
economics_small <- economics[1:100, ]

ggplot(economics_small, aes(date, unemploy)) +
  geom_line(
    linewidth = 1.5,
    color = get_raven_palette("chocolate")[3]
  ) +
  labs(
    title = "Line Plot",
    subtitle = "Unemployment over time",
    x = "Date",
    y = "Unemployed"
  ) +
  theme_raven(
    background_type = "transparent",
    font_theme = "simple",
    color_theme = "chocolate"
  )
```

# Boxplot

```{r}
ggplot(mpg, aes(class, hwy, fill = class)) +
  geom_boxplot() +
  scale_fill_manual(values = get_raven_palette("witchy")) +
  labs(
    title = "Boxplot",
    subtitle = "Highway MPG distribution",
    x = "Class",
    y = "MPG"
  ) +
  theme_raven(
    background_type = "transparent",
    font_theme = "elegant",
    color_theme = "witchy"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_boxplot(color = "#14070A")
```

# Dot plot (staked)

``` {r}
ggplot(mpg, aes(x = class, y = hwy, fill = class)) +
  geom_dotplot(
    binaxis = "y",
    stackdir = "center",
    dotsize = 0.6
  ) +
  scale_fill_manual(values = get_raven_palette("witchy")) +
  labs(
    title = "Dot Plot",
    subtitle = "Highway MPG distribution by class",
    x = "Class",
    y = "Highway MPG"
  ) +
  theme_raven(
    background_type = "vintage",
    font_theme = "simple",
    color_theme = "witchy"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Heatmap

``` {r}
corr <- cor(mtcars)

corr_df <- as.data.frame(as.table(corr))

ggplot(corr_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_gradientn(colors = get_raven_palette("vampire_heatmap")) +
  labs(
    title = "Heatmap",
    subtitle = "Correlation matrix",
    x = "",
    y = ""
  ) +
  theme_raven(
    background_type = "dark",
    font_theme = "elegant",
    color_theme = "vampire_heatmap",
    major_grid = FALSE
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Histogram

``` {r}
ggplot(diamonds, aes(carat, fill = cut)) +
  geom_histogram(bins = 30, alpha = 0.8) +
  scale_fill_manual(values = get_raven_palette("witchy")) +
  labs(
    title = "Histogram",
    subtitle = "Diamond sizes",
    x = "Carat",
    y = "Count"
  ) +
  theme_raven(
    background_type = "dark",
    font_theme = "elegant",
    color_theme = "witchy"
  )
```

# Faceted scatter

``` {r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_raven_point(dot_type = "triangle", size = 4) +
  scale_color_manual(values = get_raven_palette("lotus")) +
  facet_wrap(~class) +
  labs(
    title = "Faceted Scatter",
    subtitle = "Engine displacement vs MPG",
    x = "Displacement",
    y = "MPG"
  ) +
  theme_raven(
    background_type = "vintage",
    font_theme = "simple",
    color_theme = "lotus"
  )
```
